---
title: "DroughtNet - Community Composition Trajectory Analysis Report"
author: "NCEAS Scientific Computing Support Team"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Housekeeping

Begin by loading the packages that we will need for doing various community composition analysis and visualization. We need `tidyverse` for various wrangling tasks, `RRPP` for trajectory analysis, and Nick Lyon's development package `supportR` has a convenient NMS ordination function. We'll need `vegan` for multidimensional scaling (prerequisite to NMS).

```{r libraries, message = F, warning = F}
# install.packages("librarian")
librarian::shelf(tidyverse, RRPP, supportR, vegan)
```

We will also need to load some composition data to analyze.

```{r load-data}
all_comp <- read.csv(file.path("cover_ppt_map_10-08-2022.csv"))
```

Subset to a single site for exploratory purposes and then do some quick processing to simplify the data file for analysis

```{r subset-data}
comp <- all_comp %>%
  # Filter to one site
  dplyr::filter(site_code == "antelope.us") %>%
  # Generate a single column for block + plot + subplot
  dplyr::mutate(block_plot_subplot = paste(block, plot, subplot, sep = "_")) %>%
  # Pare down to only needed columns
  dplyr::select(year, trt, block_plot_subplot, Taxon, max_cover) %>%
  # Remove spaces from Taxon column
  dplyr::mutate(Taxon = gsub(pattern = " ", replacement = "_",
                             x = Taxon)) %>%
  # Pivot to wide format 
  tidyr::pivot_wider(names_from = "Taxon",
                     values_from = max_cover,
                     values_fill = 0)

# Glimpse it!
dplyr::glimpse(comp)
```

## perMANOVA

To begin, we want to know whether the composition of our communities differs by treatment, year, or the interaction of treatment by year. We could assess this by running a series of univariate ANOVAs where each test used a different species and then we could try to compare results across these tests to get a sense for whether our whole community differs among our treatments. Permutation Multivariate Analyses of Variance (or perMANOVAs) are a multivariate equivalent of an ANOVA that lets us more directly assess differences in community composition.

The `RRPP` package lets this happen relatively simply but requires a special dataframe format to function properly.

```{r rrpp-df}
# First, make our community composition information into a separate matrix
comm_mat <- comp %>%
  dplyr::select(-year, -trt, -block_plot_subplot) %>%
  as.matrix()

# Next, define the RRPP dataframe
comp_rdf <- RRPP::rrpp.data.frame("treatment" = comp$trt,
                                  "year" = comp$year,
                                  "plot" = comp$block_plot_subplot,
                                  "community" = comm_mat)
```

With that defined, we can run a perMANOVA with `RRPP`'s `lm.rrpp` function. It is called "lm" rather than "ANOVA" because an ANOVA is--statistically--just a special case of a linear model.

```{r rrpp-fit}
# Fit the model
mod_fit <- RRPP::lm.rrpp(community ~ treatment * year,
                         data = comp_rdf, # data = RRPP 'dataframe'
                         iter = 999, # 999 permutations
                         RRPP = TRUE # permute residuals
                         )

# Get the summary values (with F statistics)
anova(mod_fit, effect.type = "F")
```
 
Now that we have results, we can create an NMS to visualize these differences

```{r multi-dimension_scaling}
# Perform multidimensional scaling
mds_obj <- vegan::metaMDS(comp[-c(1:3)], distance = "bray",
                            autotransform = F, expand = F,
                            k = 2, try = 100)
```

Once we've done the multidimensional scaling, we can do the ordination.

```{r nms}
# Create a column for the explanatory variables of each row
comp$nms_group <- paste(comp$trt, comp$year, sep = "_")

# Identify correct number of colors / shapes
col_length <- length(unique(comp$nms_group)) / 2
ide_colors <- c(rep("mediumseagreen", times = col_length),
                rep("rosybrown4",times = col_length))
ide_lines <- c(rep(1, times = col_length),
                rep(2, times = col_length))

# Feed those vectors into the ordination function
supportR::nms_ord(mod = mds_obj, groupcol = comp$nms_group,
               shapes = rep(x = 21:25, times = 4),
               colors = ide_colors,
               lines = ide_lines,
               title = "antelope.us",
               leg_pos = 'topright')
```

## Trajectory Analysis

Trajectory analysis actually accepts the same perMANOVA fit created above.

```{r traj-fit}
traj_fit <- RRPP::trajectory.analysis(fit = mod_fit,
                                      # Groups are control vs. drought
                                      groups = comp_rdf$treatment,
                                      # Trajectory "points" are years
                                      traj.pts = comp_rdf$year)
```

Trajectory analysis gathers results for (1) the distance of the change through time, (2) the "shape" of that change, and (3) the correlations (e.g., "angles") of those differences.

```{r traj-dist}
# Magnitude distance (MD)
summary(traj_fit, attribute = "MD")
```

```{r traj-shapes}
# Shape Differences (SD)
summary(traj_fit, attribute = "SD")
```

```{r traj-angles}
# Angles (TC)
summary(traj_fit, attribute = "TC", angle.type = "deg")
```

These will hopefully make more sense in the context of a plot so let's generate that here.

```{r traj-plot-1}
# Create a plot object
traj_plot <- plot(traj_fit, pch = 21, cex = 0.7, col = "gray")

# Then add trajectories
RRPP::add.trajectories(traj_plot, traj.pch = 21)
```
