---
title: "IDE_CommAnalysis_MIA_04052022"
author: "Maggie Anderson"
date: "4/5/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Goal: sythesize plant community cover data from IDE
# Contributors: Tim Ohlert, Kate Wilkins, Maggie Anderson

# Tim's code 

### 1) setup
```{r setup, include=FALSE}
# clear workspace
  rm(list = ls()) 

knitr::opts_chunk$set(echo = TRUE)

# load packages
pacman::p_load(codyn,         # diversity metrics
               ggthemes,
               plyr,
               RColorBrewer,  # fun colors for plots
               tidyverse,     # wrangling
               )         
```

### 2) load data, summaries
```{r}
full_cover <- read.csv("~/Dropbox/IDE/data_processed/full_cover_04-04-2022.csv")

summary(full_cover)

summary(full_cover$max_cover)

hist(full_cover$max_cover) 
```

### 3) unique column values (facors)
```{r}
unique(full_cover$trt)
unique(full_cover$functional_group)
unique(full_cover$live)
unique(full_cover$ps_path)
unique(full_cover$local_lifespan)
unique(full_cover$n_treat_years)

length(unique(full_cover$Taxon))
```

### 4) subset data
```{r}
grass <- subset(full_cover, functional_group == "GRASS")
summary(grass)
graminoid <- subset(full_cover, functional_group == "GRAMINOID")
summary(graminoid)
null <- subset(full_cover, functional_group == "NULL")
summary(null)
```

### 5) make some plot level aggregate values for SR, H, %grass/forb
```{r}
# Make some plot level aggregate values
# -species richness
# -evenness
# -percent grass
# -percent forb

full_cover$functional_group <- ifelse(full_cover$functional_group == "GRASS", "GRAMINOID", full_cover$functional_group)
comb.by.plot <- full_cover%>%
  subset(live == 1)%>%
  subset(trt == "Control" | trt == "Drought")%>%
  ddply(.(site_code, block, plot, subplot, year, first_treatment_year, first_treatment_date, cover_date, n_treat_days, n_treat_years, trt),
        function(x)data.frame(
          species_richness = length(x$Taxon),
          cover = sum(x$max_cover),
        EQ = community_structure(
          x,
          time.var = NULL,
          abundance.var = "max_cover",
          replicate.var = NULL,
          metric = c("EQ"))$EQ
          
          
        ))
head(full_cover)
```

# Kate's code

### 6) limit to sites with suffucuent no. of reps (community data)
```{r}
uniq.plot<- comb.by.plot %>% 
  dplyr::filter(trt %in% c("Drought","Control"))%>%
  dplyr::select(site_code,year,plot,trt)%>%
  dplyr::distinct(site_code,year,plot,trt)%>%
  dplyr::as_tibble()

Plot.trt.ct <- uniq.plot%>% dplyr::group_by(site_code,trt,year) %>% dplyr::summarise(Plot.count=n())

#Making to wide format to see which sites do not have both treatments in a year
Plottrt_wide<-spread(Plot.trt.ct,trt,Plot.count)
Plottrt_wide[is.na(Plottrt_wide)] <- 0

#Remove sites and years that don't have both control and drought plots
  #OR that only have one rep of drought
  #brokenh.au in 2018 and 2019 (0 drought plots)
  #chacra.ar in 2016 (1 control plot) #Keeping
  #chilcasdrt.ar in 2016 and 2020 (0 drought plots)
  #charleville.au in 2017 (1 control plot) #Keeping
  #charleville.au in 2019 (1 drought plot)
  #cobar.au in 2018 (1 control plot) #keeping

Plottrt_wide1<-Plottrt_wide%>%
  dplyr::filter(Drought>=2 & Control>=1)%>%
  dplyr::as_tibble()

#Switch back to long format for merge
Plot.trt.ct2<-gather(Plottrt_wide1,trt,Plot.count,Drought:Control)

#Merge unique trt count back with data frame to filter
comb.by.plot<-merge(comb.by.plot,Plot.trt.ct2,by=c("site_code","trt","year"))
```

### 7) calculate drought severity
```{r}
# read in precip data
ppt_data<-read.csv("~/Dropbox/IDE MS_Single year extreme/Data/precip/precip_by_trmt_year_with_percentiles_2021-05-12.csv")
head(ppt_data)

# pull precip in drought plots and WorldClim values for each YEAR (maybe change this later)
ppt_data_yr <- ppt_data[,c(1,2,10:14)]

# remove duplicate site_year plots
ppt_data_yr$site_code_yr <- paste(ppt_data_yr$site_code, ppt_data_yr$year, sep = "_")
ppt_data_yr <- ppt_data_yr[!duplicated(ppt_data_yr$site_code_yr),] # remove duplicate cover dates

# drought severity calculation
ppt_data_yr$drtpct_map<-((ppt_data_yr$ppt_drought-ppt_data_yr$wc_map)/ppt_data_yr$wc_map)
  # ppt_drought = precip in drought plot
  # wc_map = MAP data from WorldClim


# merge with cover dataset to get year-specific values
#test <- left_join(full_cover1, ppt_data_yr[,c()])


# join datasets
comb.by.plot <- left_join(comb.by.plot,ppt_data_yr[,c(1:2,9)], by = c("site_code","year")) # STILL WORKING ON THIS
```

# Maggie's code

### 8) calculate richness and evenness response
```{r}
# calculate species richness response
comb.wide.sr <- comb.by.plot %>%
  pivot_wider(names_from = trt,
              values_from = species_richness) %>%
  select(site_code,year,Control,Drought) %>%
  group_by(site_code, year) %>% 
  replace(is.na(.), 0) %>%
  summarise_each(funs(sum)) %>%
  mutate(sr_response = log(Drought/Control)) 

comb.wide.sr <- left_join(comb.wide.sr,
            distinct(comb.by.plot[,c(1,3,11,15:16)]), 
            by = c("site_code","year")) # join to get n_treat_years, Plot.count, & drtpcp_map

# calculate evenness response
comb.wide.eq <- comb.by.plot %>%
  pivot_wider(names_from = trt,
              values_from = EQ) %>%
  select(site_code,year,Control,Drought) %>%
  group_by(site_code, year) %>% 
  replace(is.na(.), 0) %>%
  summarise_each(funs(sum)) %>%
  mutate(eq_response = log(Drought/Control))

comb.wide.eq <- left_join(comb.wide.eq,  
            distinct(comb.by.plot[,c(1,3,11,15:16)]), 
            by = c("site_code","year"))

# combine into the single dataframe
comb_rr <- left_join(comb.wide.sr, distinct(comb.wide.eq[,c(1:2,5)]), by = c("site_code","year"))
```

### 9) re-add site information 
```{r}
# add site_level information to the dataframe
site_info <- read.csv("~/Dropbox/IDE MS_Single year extreme/Data/Site_Elev-Disturb.csv")
head(site_info)

# join with comb_rr dataset
comb_rr <- left_join(comb_rr, distinct(site_info[,c(2:9)]), by = "site_code")
```

### 10) plots: total richness and evenness response ratio (RR) by drought severity
```{r}
comb_rr %>%
  filter(n_treat_years %in% (1:4)) %>%
ggplot(aes(x=drtpct_map, y = sr_response)) +
  geom_point(alpha = 0.5, width = 0.1, size=1) +
  geom_smooth(method = "lm") +
  theme_bw() +
  #facet_wrap(~n_treat_years) +
  theme(panel.grid.major.x = element_blank(),
         panel.grid.major.y = element_line(size=.1, color="grey")) +
  xlab("Drought Severity") +
  ylab("Species Richness Response") +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "black", size=1) 

# negative numbers: less SR in drought vs control plots

comb_rr %>%
  filter(n_treat_years %in% (1:4)) %>%
ggplot(aes(x=drtpct_map, y = eq_response)) +
  geom_point(alpha = 0.5, width = 0.1, size=1) +
  geom_smooth(method = "lm") +
  theme_bw() +
  #facet_wrap(~n_treat_years) +
  theme(panel.grid.major.x = element_blank(),
         panel.grid.major.y = element_line(size=.1, color="grey")) +
  xlab("Drought Severity") +
  ylab("Evenness Response") +
 geom_vline(xintercept = 0, linetype="dashed", 
                color = "black", size=1)
```

### separate RR out by habitat type and treatment years
```{r}
# separate each by year: species richness
comb_rr %>%
  filter(n_treat_years %in% (1:4)) %>%
  filter(habitat.type == "Grassland" | habitat.type == "Shrubland") %>%
ggplot(aes(x=drtpct_map, y = sr_response)) +
  geom_point(aes(col = habitat.type, shape = habitat.type),alpha = 0.5, width = 0.1, size=2) +
  geom_smooth(method = "lm") +
  scale_color_manual(name="Ecosystem Type",labels = c('Grassland','Shrubland'),values=c("mediumseagreen","wheat3"))+
  facet_grid(~n_treat_years) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
         panel.grid.major.y = element_line(size=.1, color="grey")) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "black", size=1) +
  xlab("Drought Severity") +
  ylab("Species Richness Response")

# separate each by year: evenness
comb_rr %>%
  filter(n_treat_years %in% (1:4)) %>%
  filter(habitat.type == "Grassland" | habitat.type == "Shrubland") %>%
ggplot(aes(x=drtpct_map, y = eq_response)) +
  geom_point(aes(col = habitat.type, shape = habitat.type),alpha = 0.5, width = 0.1, size=2) +
  geom_smooth(method = "lm") +
  scale_color_manual(name="Ecosystem Type",labels = c('Grassland','Shrubland'),values=c("mediumseagreen","wheat3"))+
  facet_grid(~n_treat_years) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
         panel.grid.major.y = element_line(size=.1, color="grey")) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "black", size=1) +
  xlab("Drought Severity") +
  ylab("Evenness Response")
```



_______________________________________

